---
title: "EFA"
author: "Anita Restrepo"
date: "`r format(Sys.time(), '%b %d, %Y')`"
output: html_document
---
  
```{r include = FALSE}
# ensures none of the code chunks are actually run (because non-functional
# code snippets will not allow for knitting)
knitr::opts_chunk$set(eval = FALSE)
```
  
Libraries required:  
```{r}
library(paran)
```

This code for exploratory factor analysis (EFA) is very badly documented (sorry but I can't remember what each line does eek) but should be a decent starting point!   
  
This was originally done on the Internet Addiction Test (IAT) and so the individual questionnaire items are labelled IAT_itemNumber. The dataframe called IAT_Items includes the item-level responses for all participants for all items of the IAT WITHOUT the Subject ID or any other variables that could confuse the EFA.    
  
This code doesn't specify how many factors to retain:  
```{r}
fa.parallel(IAT_Items, fa = "fa", cor = "poly") 
# fa specifies factor analysis instead of principal component analysis (PCA)
# poly specifies variables are categorical
```
  
This code runs an EFA retaining 3 factors:  
```{r}
fa.poly(IAT_Items, nfactors = 3)
```
  
This code runs a parallel analysis to see how many factors SHOULD be retained:  
```{r}
paran(IAT_Items,
      iterations = 10000, # for bootstrapping
      centile = 95,
      quietly = FALSE, 
      status = TRUE,
      all = TRUE,
      cfa = TRUE,
      graph = TRUE,
      color = TRUE, 
      col = c("black", "red", "blue"),
      lty = c(1, 2, 3),
      lwd = 1, 
      legend = TRUE, 
      file = "", 
      width = 640, 
      height = 640, 
      grdevice = "png", 
      seed = 0)
```
  
This also produces a scree plot (I think).  
  
This code seems to be another way to run the parallel analysis (?):  
```{r}
# run a principal components analysis
PCA <- princomp(IAT_Items)

# generate a scree plot
plot(PCA, main = "Scree Plot") 

# based on the scree plot, run the EFA retaining 7 factors
# with varimax rotation
factanal(IAT_Items, factors = 7, rotation = "varimax") 
```

