---
title: "Merge Multiple Data Frames"
author: "Anita Restrepo"
date: "`r format(Sys.time(), '%b %d, %Y')`"
output: html_document
---
  
  
 
  
```{r include = FALSE}
# ensures none of the code chunks are actually run (because non-functional
# code snippets will not allow for knitting)
knitr::opts_chunk$set(eval = FALSE)
```
  
Libraries required:  
```{r}
library(plyr)
```


From Lindsay Alexander  
Merge multiple dataframes (or subsets of them) by a common variable    
  
```{r}
## create Merge function
Merge <- function(x, y){
  df <- merge(x, y, by= "common variable", all.x=TRUE, all.y=TRUE)
  return(df)
}
## use Reduce to apply Merge function (which only combines two elements (i.e. 2 dfs in this case)) to multiple elements
Merged_df <- Reduce(Merge, list(df1, df2, df3, df4, ...))
```

[About the Reduce function](https://blog.zhaw.ch/datascience/r-reduce-applys-lesser-known-brother/):  
Applies a function (in this case the Merge function we created) to a list of elements in a stepwise process: starts with the first two elements, then does the result of those two with the third element, and so on  
can set accumulate == TRUE to return not just the final result (in this case the total merge of all dfs) but also all intermediate results (the merge of dfs1 + 2, + 3, etc.)  
  
can also use the join_all() function from plyr package  
[documentation](https://www.rdocumentation.org/packages/plyr/versions/1.8.6/topics/join_all)
```{r}
join_all(dfs, by = NULL, type = "left", match = "all")
    # by = "commmon column"
    # type = type of join (left, right, inner, full
    # match = how to deal with duplicates in key ("first": match only first occurrence; "all": match all occurrences)
```

