---
title: "Probe Model Interactions with emmeans"
author: "Anita Restrepo"
date: "`r format(Sys.time(), '%b %d, %Y')`"
output: html_document
---
  
  
This code is courtesy of the wonderful Karen Smith. It uses the emmeans package to probe 2- and 3-way interactions in models, mainly HLMs. There are other things that could be done with this pakage but these are some of the basics needed for reporting.    
  
```{r include = FALSE}
# ensures none of the code chunks are actually run (because non-functional
# code snippets will not allow for knitting)
knitr::opts_chunk$set(eval = FALSE)
```
  
Libraries required:  
```{r}
library(nlme4)
library(emmeans)
```
  
# The Model  
The model with the interactions (in this case a 2-level HLM run with the lme4 package):  
  
The cyberball_condition variable is categorical with three levels.  
The cyber_session variable only has 5 values (1-5) but is being treated as numeric here.  
```{r}
model <- lmer(cyber_total ~ cyberball_condition*cyber_session*UCLA_score + (1 + cyber_session|Subject), data = dat, REML = FALSE)
```
  
# Compare Intercepts  
To run contrasts between the different levels of the categorical cyberball_condition comparing whether the **intercepts** for each condition are the same:   
  
```{r}
pairs(emmeans(model, "cyberball_condition"), adjust = NULL)
```
  
# Probe 2-way interaction  
To run contrasts between the different levels of the categorical cyberball_condition comparing whether the **slopes** (i.e. the effect of cyber_session) differs between the levels:  
  
```{r}
emtrends(model, pairwise~cyberball_condition, var = "cyber_session")
```
Note: the p-values in this output determine whether the conditions are significantly different from each other, *not* whether the slopes are significantly different from zero.  
  
To see whether the estimates (i.e. the betas) for the slopes are significantly different from zero:  
```{r}
test_betas <- emtrends(model, ~cyberball_condition, var = "cyber_session", 
         at=c(list(cyber_session = seq(1, 5))), options = list())
summary(test_betas, infer=c(TRUE,TRUE), null=0)
```
Note: here, we include `seq(1, 5)` because we are testing the slopes across the full range of values for the cyber_session variable.  
  
# Probe 3-way interaction  
Because we can't run contrasts between the conditions, the only thing we can do (apart from plotting, v important) is run significance tests for the betas at multiple combinations of values for the interacting variables.  
```{r}
UCLA.high <- M + SD
UCLA.low <- M - SD

trends <- emtrends(model, ~cyberball_condition*UCLA_score, var = "cyber_session", 
         at=c(list(cyber_session = seq(1, 5)), list(UCLA_score = c(UCLA.low, UCLA.high))), options = list())
summary(trends, infer=c(TRUE,TRUE), null=0)
```
Note: We use the full range (i.e. `seq(1, 5)`) for the cyber_session variable in order to get the slopes (i.e. trends) and +/- 1 SD for the UCLA variable, which is numerical and continuous. Because cyberball_condition is categorical (i.e. coded as a factor), emtrends will compute a beta estimate for each condition at both +1SD and -1SD for the UCLA variable and run a significance test on each estimate.  
  


  
