---
title: "Regression Tables"
author: "Anita Restrepo"
date: "8/12/2021"
output: html_document
---
```{r include = FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

This code is meant to automate the process of running regressions and pulling out relevant stats into tables to then be able to move them out of R easily.  
  
Run multiple regressions on various variables of interest within a dataframe and return a results table in csv format
```{r}
## create a list that includes the column numbers for the relevant variables to subsequently put into regression loops
Variable_Column_numbers <- c(1, 2, 3, 4, 5)

## input these variables into a loop that runs a regression of choice on each of the variables
# lapply applies my self-defined function onto each item in the list
# the function in this case is a glm (logistic regression)
# each item in my list of variables (defined as x) is being included as a predictor in the models by selecting the relevant column from my dataframe
my_regressions <- lapply(Variable_Column_numbers, function(x) glm(Dependent_Variable ~ dataframe[,x] + covariate1 + covariate2, family = binomial, data = dataframe))

## pull out the information that will go into the tables of results using multiple loops that iterate over the output of the previous regression loop
# lapply and sapply do the same thing (apply function to list) but sapply returns a vector while lapply returns a list
# outputs rounded to 2 digits
# outputs of regressions are being indexed according to value of interest (e.g. 2nd item in coefficients)
# here getting unstandardized regression coefficients, confidence intervals and p values
my_summaries <- lapply(my_regressions, summary)
my_coefs <- sapply(my_regressions, function(x) round(exp(x$coefficients[2]), digits = 2))
my_CI_1 <- sapply(my_regressions, function(x) round(exp(confint(x)[2, 1]), digits =2)) 
my_CI_2 <- sapply(my_regressions, function(x) round(exp(confint(x)[2, 2]), digits = 2))
my_ps <- sapply(my_summaries, function(x) round(x$coefficients[2, 4], digits=2))

## combine all relevant information from whole loop of regressions and save into one dataframe
# generates table with columns for relevant information (B, CIs, p values) and each row is for one variable of interest (here as predictor in regression models)
# saves table as csv file in a subfolder (Tables) within the current directory
Result_table_name <- data.frame(my_coefs, my_CI_1, my_CI_2, my_ps, row.names = NULL)
colnames(Result_table_name) <- c("B", "2.5% CI", "97.5% CI", "p value")
rownames(Result_table_name) <- c("Predictor 1", "Predictor 2", "Predictor 3")
write.csv(Result_table_name, file = "Tables/Result_table_name.csv")
```
  
Update 2/23/22: The above code doesn't capitalize on the tidyverse and is much dirtier than one would like so here is some nicer, tidyer code to achieve similar results. It won't produce the same tables but is more easily modified to suit more flexible needs.  
  
For this code below, the function was made to simultaneously loop through two sets of variables to include a corresponding pair of variables as predictors in the same model. Can be modified to only loop through one set of variables or more than two.  
  
Variable explanations:  
- dat: the dataframe where the data lives  
- items: one set of predictors (i.e. column names) to loop through  
- intensities: a corresponding set of predictors to loop through (must be the same length as items)  
- questionnaire: a single outcome variable for the regression models. Needs to be inputted into the function in the form of a string (e.g. "CESD_Total")  
```{r}
reg_by_item <- function(dat, items, intensities, questionnaire) {
  regs <- data.frame() # create empty dataframe to populate inside for loop
  for (i in 1:length(items)) { # loop through the number of elements in the "items" vector
    red_df <- dat %>% select(questionnaire, items[i], intensities[i]) # create a reduced dataframe (red_df) that only contains our variables of interest
    red_df$endorsal <- ifelse(red_df[,2] > 1, 1, 0) # optional and specific to this data (creating new variable)
    DV <- questionnaire # for inclusion in final dataframe
    mod <- summary(lm(red_df[,1] ~ red_df[,4]*red_df[,3])) # run the model and call summary to extract stats
    reg <- data.frame(DV = c(DV, DV, DV, DV), # include the DV as many times as number of coefficients in your model output
                      coefficient = c("Intercept", items[i], intensities[i], "interaction"), # label for each coefficient
                      mod$coefficients, row.names = NULL) # extract betas, standard errors, t-vals and p-vals
    regs <- rbind(regs, reg) # populate the output dataframe
  }
return(regs)  
}

# vector to pass into function for the "items" argument
bleep <- c("ACES_1", "ACES_2", "ACES_3", "ACES_4", "ACES_5", "ACES_6", "ACES_7", "ACES_8", "ACES_9", "ACES_10")
# vector to pass into function for the "intensities" argument
bloop <- c("ACES_I_1", "ACES_I_2", "ACES_I_3", "ACES_I_4", "ACES_I_5", "ACES_I_6", "ACES_I_7", "ACES_I_8", "ACES_I_9", "ACES_I_10")

# create three dataframes containing the outputs for the regressions for three deffierent DVs
Lon_CTQ <- reg_by_item(df, bleep, bloop, "Lon_Total")
CESD_CTQ <- reg_by_item(df, bleep, bloop, "CESD_Total")
PSS_CTQ <- reg_by_item(df, bleep, bloop, "PSS_Total")

```
